---
title: "Cookbook - (1) Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cookbook_overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(pacta.multi.loanbook)

plot_table <- function(table) {
  table_plot <- gt::gt(dplyr::select(table, -"dataset"))
  
  table_plot <- 
    gt::cols_width(
      .data = table_plot,
      column ~ gt::px(150),
      typeof ~ gt::px(90)
    )
  
  table_plot <-
    gt::tab_style(
      data = table_plot,
      style = gt::cell_text(size = "smaller"),
      locations = gt::cells_body(columns = 1:2)
    )
  
  table_plot <-
    gt::tab_options(
      data = table_plot,
      ihtml.active = TRUE,
      ihtml.use_pagination = FALSE,
      ihtml.use_sorting = TRUE,
      ihtml.use_highlight = TRUE
    )
  
  gt::fmt_passthrough(table_plot)
}
```
# Overview

This cookbook provides a step-by-step guide to running the PACTA for Supervisors analysis using the `pacta.multi.loanbook` package. The analysis is designed to help financial supervisors assess the alignment of banks' loan books with the Paris Agreement goals.

## What is the PACTA for Supervisors analysis?

PACTA for Supervisors is based on the PACTA methodology, which assesses the alignment of financial portfolios with climate goals utilizing forward-looking asset-based company data (ABCD) that is linked to financial assets and compares the production profiles of those companies with technology and emissions pathways from climate transition scenarios at the sector and/or technology level.

## Who is this tool built for?

The PACTA for Supervisors analysis is primarily designed to be run by financial supervisors on their own, with minimal additional guidance. However, it can also be a useful tool for any other user, in case they would like to run a PACTA analysis on multiple loan books.

The main difference between this tool and the individual PACTA for Banks software packages is that this tool aims to facilitate the analysis of multiple loan books at once, streamlining the process as much as possible, to keep the burden for the user to a minimum. As such it is a helpful tool for anyone who would like to run a PACTA analysis on multiple loan books.

## What can the results of the PACTA for Supervisors analysis be used for?

Financial supervisors or regulators can use the results of the analysis to assess the alignment of banks' loan books with the Paris Agreement goals, to identify sectors where banks may need to take action to improve alignment, and to screen the financial system for potential climate-related transition risks. The analysis can be parameterised in different ways to explore patterns across the analysed loan books. If run by private institutions, the results may additionally be used to identify opportunities for climate-aligned investments, as well as for detecting individual counterparties that may be exposed to climate-related transition risks and may therefore require dedicated focus in the risk management process.

Users will be able to obtain both tabular outputs and plots that can be used for any of the above use cases. The tabular output further enables processing of the alignment results in other models or tools, for example as an input into financial risk models, or as a recurring input into internal monitoring systems.

The level of granularity of the outputs allows for a systematic analysis of climate alignment starting at the financial system level (across all analyzed loan books together), down to the individual counterparty level, and across sectors and technologies. Grouping of the results at any of these levels by additional dimensions can easily be achieved using the configuration file.

## What are the main steps of the analysis?

The main steps of the analysis are as follows:

1. Data preparation: Prepare the ABCD data and optionally a custom sector split.
2. Matching process: Match the raw loan books to the ABCD data.
3. Prioritization of loan books; Match success and coverage diagnostics: Prioritize the matched loan books and analyze their coverage.
4. Run PACTA for Supervisors analysis: Run the analysis based on the parameters set in the `config.yml` file.

This cookbook will guide you through each of the steps of the analysis in detail, explain the required input data sets and software, and provide guidance on how to interpret the results.

